Learning from Massive Noisy Labeled Data for Image Classification

### 2015-04-13 paper.tex
Update paper.tex
Modify the introduction and related work

Deep learning from large-scale supervised training dataset has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2014panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect large amount of annotations from the Internet web images~\cite{gong2014multiview} (\ie extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to various types of noise (\eg labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various label noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise~\cite{bartlett2006convexity,teng2001comparison}. Other data cleansing algorithms are proposed~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people initialize parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2014learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. Since different sellers have their own bias on different categories, they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or the false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model builds the relations among images, noisy labels, ground truth labels, and noise types, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

Our contribution comes from three aspects. First, we study the cause of noisy labels in real-world data and describe it with a novel probabilistic model. Second, we integrate the model into a deep learning framework to help detect and correct wrong labels. We study different training strategies to make the CNNs learn from better supervisions. Finally, we collect a large-scale clothing dataset with both noisy and clean labels, which will be released for academic use.

@@

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including the effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2014classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many works have shown that label noise can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However, Bartlett \etal~\cite{bartlett2006convexity} proved that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} showed that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2014classification}. On the other hand, some label noise cleansing methods were proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches were difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they might remove too many instances and the overcleansing could reduce the performances of classifiers~\cite{matic1992computer}.

>>>>

Deep learning with large-scale supervised training dataset has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2014panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect a large amount of annotations from the Internet web images~\cite{gong2014multiview} (\ie extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to various types of noise (\eg labeling mistakes from annotators or computing errors from extraction algorithms). Many studies have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various label noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise~\cite{bartlett2006convexity,teng2001comparison}. Other data cleansing algorithms are proposed~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people initialize parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2014learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using the huge amount of noisy labeled data.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modifies a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumes noisy labels are conditionally independent of input images given clean labels. However, when examining our collected dataset, we find that this assumption is too strong to fit the real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. Since different sellers have their own bias on different categories, they may provide wrong keywords for similar clothes. We observe these visual patterns and hypothesize that they are important to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or the false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

Our proposed probabilistic model captures the relations among images, noisy labels, ground truth labels, and noise types, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on our real-world clothing dataset indicate that our model can better detect and correct the noisy labels.

Our contributions fall in three aspects. First, we study the cause of noisy labels in real-world data and describe it with a novel probabilistic model. Second, we integrate the model into a deep learning framework and explore different training strategies to make the CNNs learn from better supervisions. Finally, we collect a large-scale clothing dataset with both noisy and clean labels, which will be released for academic use.

@@

\section{Related Work} % (fold)
\label{sec:related_work}
Learning with noisy labeled training data has been extensively studied in the machine learning and computer vision literature. For most of the related work including the effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2014classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many studies have shown that label noises can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However, Bartlett \etal~\cite{bartlett2006convexity} prove that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} show that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2014classification}. On the other hand, some label noise cleansing methods were proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches were difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they might remove too many instances and the overcleansing could reduce the performances of classifiers~\cite{matic1992computer}.
